---
title: "AvocadoPrices"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Predicting Avocado Prices

Link to Kaggle dataset: https://www.kaggle.com/neuromusic/avocado-prices
Predict: AveragePrice

Methods: Weighted Least Squares

Import the data

```{r}
raw_avocado = read.csv("C:/Users/peter/OneDrive/Documents/avocado.csv")
# 18,249
```

Chgeck

```{r}
hist(raw_avocado$AveragePrice)
#sum(raw_avocado$AveragePrice > 2.6)
```



Remove any nulls in the data

```{r}
raw_avocado = na.omit(raw_avocado)
```

Rename those three columns: https://loveonetoday.com/how-to/identify-hass-avocados/

Total Volume - Total sales volume of avocados
4046 - Total sales volume of avocados with PLU 4046
4225 - Total sales volume of avocados with PLU 4225
4770 - Total sales volume of avocados with PLU 4770

```{r}

```

Extracting month up from the Date column and then deleting date

```{r}
# Month
raw_avocado$month = substr(raw_avocado$Date, 6, 7)

# Drop Old Date Column and New date column but leave month and year
raw_avocado$Date = NULL
```

Eliminate Column "X"

```{r}
raw_avocado$X <- NULL
```

Look at unique values for each column

```{r}
#unique(raw_avocado$XLarge.Bags) # continuous variable, numeric
# unique(raw_avocado$type) # factor level
unique(raw_avocado$region) # factor level 
#unique(raw_avocado$YEAR) # factor level 
```
Factor Level variables 
1. Type (conventional, organic)
2. Region (54 regions) Should you narrow these down into 4 or 5?
3. Year (2015-2018)

Trim Region down into 5 groups: NE, SE, MW (MidWest), W (West not in CA), CA

Unknown if Portland refers to OR or ME?? Will omit from data

```{r}
CA <- c("California","LosAngeles","Sacramento","SanDiego","SanFrancisco")

W <- c("Boise","DallasFtWorth","Denver","LasVegas","PhoenixTucson","Seattle","Spokane","West","WestTexNewMexico")

SE <- c("BaltimoreWashington","Charlotte","Houston","Jacksonville","Louisville","MiamiFtLauderdale","Midsouth","Nashville","NewOrleansMobile","Orlando","RaleighGreensboro","RichmondNorfolk","Roanoke","SouthCarolina","SouthCentral","Southeast","Tampa")

NE <- c("Albany","Boston","BuffaloRochester","HarrisburgScranton","HartfordSpringfield","NewYork","Northeast","NorthernNewEngland","Philadelphia","Pittsburgh","Syracuse")

MW <- c("Chicago","CincinnatiDayton","Columbus","Detroit","GrandRapids","GreatLakes","Indianapolis","Midsouth","Plains","StLouis")
```

Create new Region based on 5 categories created above

```{r}
raw_avocado$NewRegion = rep(NA, length(raw_avocado$region))
raw_avocado$NewRegion[which(raw_avocado$region %in% NE)] = "NE"
raw_avocado$NewRegion[which(raw_avocado$region %in% MW)] = "MW"
raw_avocado$NewRegion[which(raw_avocado$region %in% SE)] = "SE"
raw_avocado$NewRegion[which(raw_avocado$region %in% W)] = "W"
raw_avocado$NewRegion[which(raw_avocado$region %in% CA)] = "CA"

# drop region column
raw_avocado$region = NULL

# drop any rows NA
raw_avocado = na.omit(raw_avocado) # lose ~ 1k
```


Find data types for each variable

```{r}
sapply(raw_avocado,class)
```

Tell R to treat the following variables as factors

```{r}
raw_avocado$type<-as.factor(raw_avocado$type)
raw_avocado$month<-as.factor(raw_avocado$month)
raw_avocado$year<-as.factor(raw_avocado$year)
raw_avocado$NewRegion<-as.factor(raw_avocado$NewRegion)
```

Check to see if any numeric values contain zero
X4046
X4225
X4770
Total.Bags
Small.Bags
Large.Bags
XLarge.Bags 

```{r}
summary(raw_avocado)
```

Check to see which variables could benefit from log transformation

```{r}
hist(raw_avocado$AveragePrice)
hist(raw_avocado$Total.Volume)
hist(raw_avocado$X4046)
hist(raw_avocado$X4225)
hist(raw_avocado$X4770)
hist(raw_avocado$Total.Bags)
hist(raw_avocado$Small.Bags)
hist(raw_avocado$Large.Bags)
hist(raw_avocado$XLarge.Bags)
```

Pretty much all of them. Scale all numeric variables. Columns that contain '0' will be log transformed with +1

```{r}
raw_avocado$AveragePrice<-log(raw_avocado$AveragePrice)
raw_avocado$Total.Volume<-log(raw_avocado$Total.Volume)
raw_avocado$X4046<-log(raw_avocado$X4046+1)
raw_avocado$X4225<-log(raw_avocado$X4225+1)
raw_avocado$X4770<-log(raw_avocado$X4770+1)
raw_avocado$Total.Bags<-log(raw_avocado$Total.Bags+1)
raw_avocado$Small.Bags<-log(raw_avocado$Small.Bags+1)
raw_avocado$Large.Bags<-log(raw_avocado$Large.Bags+1)
raw_avocado$XLarge.Bags<-log(raw_avocado$XLarge.Bags+1)
```

Check Hist of them now

```{r}
boxplot(raw_avocado$AveragePrice)
boxplot(raw_avocado$Total.Volume)
boxplot(raw_avocado$X4046)
boxplot(raw_avocado$X4225)
boxplot(raw_avocado$X4770)
boxplot(raw_avocado$Total.Bags)
boxplot(raw_avocado$Small.Bags)
boxplot(raw_avocado$Large.Bags)
boxplot(raw_avocado$XLarge.Bags)
```

Try lm model

```{r}
fit = lm(AveragePrice ~., data=raw_avocado)
summary(fit)
par(mfrow = c(2,2))
plot(fit)
```

#######################################################################################

Linear Regression from MIDTERM

```{r}
set.seed(1,sample.kind = "Rounding")
library(leaps)
model_1.regfit<-regsubsets(AveragePrice~.,data=raw_avocado,nvmax=75)
model_1.regfit.summary=summary(model_1.regfit)
```

```{r}
par(mfrow=c(2,2))
plot(model_1.regfit.summary$rss,xlab="Number of Variables",ylab="RSS",type="l")
plot(model_1.regfit.summary$adjr2,xlab="Number of Variables",ylab="Adj R2",type="l")
plot(model_1.regfit.summary$bic,xlab="Number of Variables",ylab="BIC",type="l")
plot(model_1.regfit.summary$cp,xlab="Number of Variables",ylab="CP",type="l")
```
```{r}
# Plot adjusted r2
plot(model_1.regfit,scale="adjr2")

# store regfit into variable to view summary
model_1.regfit.summary<-summary(model_1.regfit)
model_1.regfit.summary$adjr2

# adjr
which.max(model_1.regfit.summary$adjr2)
```


#########################################################################################

Method:
Iteratively Weighted Least Squares Regression

# Don't use
```{r}
fit.w = lm(AveragePrice ~., data = raw_avocado)

oldcoef = rep(0,length(fit.w$coef))
newcoef = fit.w$coef
iter = 0

while(sum(abs(oldcoef)) > 0.0001 & iter < 100){
  w = 1/(fit.w$fittedvalues^2) # find new weights
  fit.w = lm(AveragePrice ~., data = raw_avocado, weights = w) # Re-fit the model
  
  iter = iter + 1
  oldcoef = newcoef
  newcoef = fit.w$coef
}
plot(fit.w)
iter
```



```{r}
fit.w=lm(AveragePrice ~ Total.Volume + X4046 + X4225 + X4770 + Total.Bags + Small.Bags + Large.Bags + type + year + month + NewRegion, data = raw_avocado)

oldcoef=rep(0,length(fit.w$coef))
newcoef=fit.w$coef
iter=0

while(sum(abs(oldcoef-newcoef))>.0001 & iter < 100){
MAR=median(abs(fit.w$residuals))
k=1.345*MAR/0.6745
w=pmin(k/abs(fit.w$residuals),1)
fit.w=lm(AveragePrice ~ Total.Volume + X4046 + X4225 + X4770 + Total.Bags + Small.Bags + Large.Bags + type + year + month + NewRegion, data = raw_avocado, weights=w)

iter=iter+1
oldcoef=newcoef
newcoef=fit.w$coef
}
iter
coef(fit.w)

```


Use rlm() to fit a robust regression model with Tukey’s bisquare weights

```{r}
library(MASS)
tukey.fit.bisquare=rlm(AveragePrice ~ Total.Volume + X4046 + X4225 + X4770 + Total.Bags + Small.Bags + Large.Bags + type + year + month + NewRegion, data = raw_avocado, psi = psi.bisquare)
tukey.fit.bisquare
```

Plot weights

```{r}
plot(tukey.fit.bisquare$w,las=1,cex.axis=1.2,ylab="Weights")

smallweights=which(tukey.fit.bisquare$w <.8)
showLabels(1:dim(crime)[1],tukey.fit.bisquare$w,raw_avocado$AveragePrice,method=smallweights)

```


########################################################################################

Robust Regression CV
This works!

```{r}
library(MASS)
# 10 fold cross validation
n=nrow(raw_avocado)
k=10
groups=c(rep(1:k,floor(n/k)),1:(n-floor(n/k)*k))

set.seed(1,sample.kind = "Rounding")
cvgroups=sample(groups,n)
bisquare.pred=rep(0,n)
huber.pred=rep(0,n)
lm.pred=rep(0,n)

for(i in 1:k){
  groupi=(cvgroups==i)
  # tukey bisquare
  fit.bisquare=rlm(AveragePrice ~ Total.Volume + X4046 + X4225 + X4770 + Total.Bags + Small.Bags + Large.Bags + type + year + month + NewRegion, data = raw_avocado[!groupi,], psi = psi.bisquare)
  bisquare.pred[groupi]=predict(fit.bisquare,newdata=raw_avocado[groupi,])
  # Huber
  fit.huber=rlm(AveragePrice ~ Total.Volume + X4046 + X4225 + X4770 + Total.Bags + Small.Bags + Large.Bags + type + year + month + NewRegion, data = raw_avocado[!groupi,], psi = psi.huber)
  huber.pred[groupi]=predict(fit.huber,newdata=raw_avocado[groupi,])
  # lm
  lm=lm(AveragePrice ~ ., data= raw_avocado[!groupi,])
  lm.pred[groupi]=predict(lm,newdata=raw_avocado[groupi,])
}
mean((bisquare.pred-raw_avocado$AveragePrice)^2)
mean((huber.pred-raw_avocado$AveragePrice)^2)
mean((lm.pred-raw_avocado$AveragePrice)^2)
# This works now!

```

###########################################################################

Method 2: Try trees

```{r}
set.seed(1)
train = sample(1:dim(raw_avocado)[1], 10000, replace=F)

library(tree)
mytree = tree(AveragePrice ~ ., data= raw_avocado[train,])
summary(mytree)


```
Try decision tree with CV

```{r}

```

```{r}
plot(mytree)
text(mytree, pretty=0)

```

```{r}
library(randomForest)
library(gbm)

#raw_avocado=raw_avocado[sample(nrow(raw_avocado),1000,replace=F),]

# 10 fold cross validation
n=nrow(raw_avocado)
k=10
groups=c(rep(1:k,floor(n/k)),1:(n-floor(n/k)*k))

set.seed(1,sample.kind = "Rounding")
cvgroups=sample(groups,n)
boost.predict=rep(-1,n)
lm.pred=rep(0,n)

for(i in 1:k){
groupi=(cvgroups==i)
# boost
boost=gbm(AveragePrice ~ ., data= raw_avocado[!groupi,],distribution = "gaussian",n.trees=1000,shrinkage=.001,interaction.depth = 3)
boost.predict[groupi]=predict(boost,newdata=raw_avocado[groupi,],n.trees=1000,type="response")
# lm
lm=lm(AveragePrice ~ ., data= raw_avocado[!groupi,])
lm.pred[groupi]=predict(lm,newdata=raw_avocado[groupi,])
}

mean((boost.predict-raw_avocado$AveragePrice)^2)
mean((lm.pred-raw_avocado$AveragePrice)^2)

```
It worked! Takes about a minute to run but it worked!


Try bagging/decision tree
WARNING: This takes FOREVER!!!

```{r}
# 10 fold cross validation
library(randomForest)

# 10 fold cross validation
n=nrow(raw_avocado)
k=10
groups=c(rep(1:k,floor(n/k)),1:(n-floor(n/k)*k))

set.seed(1,sample.kind = "Rounding")
cvgroups=sample(groups,n)
rf.predict=rep(-1,n)
bag.predict=rep(-1,n)

for(i in 1:k){
  groupi=(cvgroups==i)
# bagging
  bag=randomForest(AveragePrice ~ ., data= raw_avocado[!groupi,],mtry=12,importance=T)
  bag.predict[groupi]=predict(bag,newdata=raw_avocado[groupi,])
# random forest
  rf=randomForest(AveragePrice ~ ., data = raw_avocado[!groupi,], ntree=100, mtry=4,importance=T)
  rf.predict[groupi] = predict(rf,newdata=raw_avocado[groupi,])
}

mean((bag.predict-raw_avocado$AveragePrice)^2)
mean((rf.predict-raw_avocado$AveragePrice)^2) # 0.007644937

```





